Classless Java: Tuning Java Interfaces

### 2015-12-06 paper/sections/RelatedWork.tex
Reviewd up to 8. Last part of 8 have notes we need to discuss

\subsection{Meta-programming Competes with Language Extensions}
The most obvious solution to adding features to a language is language
extension. It is often implemented as syntactic extensions that can be desugared
to the base language. For example, the Scala compiler was extended to directly
supports XML syntax. However, this approach does not support combining multiple
extensions into one. Scala-Virtualized is an extension to Scala which allows to
blend shallow and deep embedding of DSLs. It redefines Scala structures to
method calls, which can be overridden by DSL implementors. Therefore
Scala-Virtualized also reinterprets syntax. These extensions are admittedly
useful, however they are essentially creating a fork of the language, and rarely
the new fork gains enough traction to become the main language release. % On this
% topic we mention
SugarJ~\cite{erdweg11sugarj} is a Java-based extensible language allowing
programmers to extend it with custom features by definitions in meta-DSLs (SDF,
Stratego, etc). Using SugarJ requires learning various DSLs while our approach
is just using Java. However its intention is different from ours: one of the
purpose of SugarJ is to allow easy defition of sugars while our approach aims to
implement the desired program transformation without just in Java.

@@

Java-like languages tend to sit in the middle of two extremes: libraries
can not influence the type system, so many solutions valid in python or other
languages are not be applicable, or may be applicable at the cost of losing
safety.

@@

Other languages offer even stronger support to safe code manipulation:
Template Haskell~\cite{sheard2002template}, F\# (type providers)~\footnote{http://research.microsoft.com/fsharp/} and MetaFjig (Active Libraries)~\cite{servetto2010metafjig}
all allow to execute code at compile time and to generate code/classes that are transparently integrated in the program that is being generated/processed/compiled.
In particular, MetaFjig offers a property called \emph{meta-level-soundness}. In short this property ensures by construction that library code (even if wrong or non nonsensical) would never generate ill-typed code. This is roughly equivalent to what we state and manually proof in Lemma 2 for our particular transformation.
Since MetaFjig is not working on annotated classes, there is no direct equivalence on the overall theorem of safety we shown.

>>>>

\subsection{Meta-programming Competes with Language Extensions}
The most obvious solution to adding features to a language is language
extension. It is often implemented as syntactic extensions that can be desugared
to the base language. For example, the Scala compiler was extended to directly
supports XML syntax. However, this approach does not support combining multiple
extensions into one. Scala-Virtualized is an extension to Scala which allows to
blend shallow and deep embedding of DSLs. It redefines Scala structures to
method calls, which can be overridden by DSL implementors. Therefore
Scala-Virtualized also reinterprets syntax. These extensions are admittedly
useful, however they are essentially creating a fork of the language, and rarely
the new fork gains enough traction to become the main language release. % On this
% topic we mention
SugarJ~\cite{erdweg11sugarj} is a Java-based extensible language allowing
programmers to extend it with custom features by definitions in meta-DSLs (SDF,
Stratego, etc). Using SugarJ requires learning various DSLs while our approach
is just using Java. However its intention is different from ours: one of the
purpose of SugarJ is to allow easy defition of sugars while our approach aims to
implement the desired program transformation in pure Java.

@@

Java-like languages tend to sit in the middle of two extremes: libraries
can not influence the type system, so many solutions valid in python or other dynamic
languages are not be applicable, or may be applicable at the cost of losing
safety.

@@

Other languages offer even stronger support to safe code manipulation:
Template Haskell~\cite{sheard2002template}, F\# (type providers)~\footnote{http://research.microsoft.com/fsharp/} and MetaFjig (Active Libraries)~\cite{servetto2010metafjig}
all allow to execute code at compile time and to generate code/classes that are transparently integrated in the program that is being generated/processed/compiled.
In particular, MetaFjig offers a property called \emph{meta-level-soundness}. In short this property ensures by construction that library code (even if wrong or non nonsensical) would never generate ill-typed code. This is roughly equivalent to 
\textit{Self coherence}, that we manually proof in the Appendix.
Since MetaFjig is not working on annotated classes, there is no direct equivalence on the overall theorem of safety we shown.
